
In many cases (see Section~\ref{sec:xml:usecases}) an ASCII representation of a
\nexus-file is required. There are two principle directions an ASCII
representation of a \nexus-file can be used
\begin{description}
    \item[From \nexus\ to ASCII]  an existing \nexus-file is converted to 
        its ASCII representation
    \item[From ASCII to \nexus] a new \nexus-file is created based on the 
        information provided by an ASCII represenation of the file.
\end{description}

%%%===========================================================================
\section{Use cases}\label{sec:ascii:usecases}

\subsection{Generating \nexus\ structures from XML}

\nexus-files, when properly used, can store much more information than a typical
ASCII data file.
The structure of a \nexus\ file can become rather complex. In a rather complex 
environment like at a synchrotron beamline it would be naive to use static 
code to create the file. A more feasible approach would be to implement 
generic code which generates the structure of a file from an XML template. 
The template could be generated either manually by the user or by means 
of another program written in whatever language is reasonable for this purpose. 

\subsection{Metadata ingestion of a file}

In some situations a $3$rd party may needs to process some of the metadata 
stored in the file while not having access to the file. One possible application
would be a data catalogue. Instead of making the file available the XML dump 
of the \nexus\ file could be sent to the system while moving the \nexus\ file 
to its final location on the storage system.

%%%===========================================================================
\section{Necessary limitations of an ASCII representation}
\label{sec:ascii:limitations}

The ASCII representation of a \nexus-file is considered as a utility to provide
a solution to the above use cases. Thus, every ASCII representation must have 
limitations. In particular a certain asymmetry exists between the ASCII 
representation obtained by one of the two basic use cases. 
The ASCII representation produced by a \nexus\ to ASCII converter is not
necessarily  exactly the same as the one used to create the file in the first
place.
\begin{figure}[tb]
\centering
\begin{tikzpicture}
    \node[artifact] (input file) {Input ASCII representation};
    \node[process,right  = of input file] (ascii to nexus) { 
    ASCII to \nexus\ conversion};
    \node[artifact,right = of ascii to nexus] (nexus file) 
    {\nexus-file};
    \node[process,right = of nexus file] (nexus to ascii) {
    \nexus\ to ASCII conversion};
\end{tikzpicture}
\end{figure}

%%%===========================================================================
\section{\nexus\ and XML}
\input{tex/xml.tex}
